---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setoptions, echo=FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

```{r setup, echo=FALSE}
library(ggplot2)
library(dplyr)
library(xtable)

```


## Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv")
activity <- transform(activity, date = as.Date(activity$date, "%Y-%m-%d"))
```

## What is mean total number of steps taken per day?

```{r}
steps <- activity %>% group_by(date) %>% summarise(Totalsteps = sum(steps), Mean = mean(steps), Median = median(steps))

```

### Summary of total, mean and median of steps taken by each day

```{r, results='asis'}
stepsxt <- transform(steps, date = as.character(steps$date))
xt <- xtable(stepsxt)
print(xt, type = "html", include.rownames = FALSE)
```

### Histgram of total number of steps

```{r}
             
barplot(steps$Totalsteps, space = 0, names.arg = steps$date,
        xlab = "Date",
        ylab = "Number of steps",
        main = "Total number of steps")

```


## What is the average daily activity pattern?

### Plot of average daily activity pattern 
```{r}
interval <- activity %>% group_by(interval) %>% summarise(Mean = mean(steps, na.rm = TRUE))

plot(interval, type = "l",
     xlab = "Interval",
     ylab = "Number of steps",
     main = "Average daily activity pattern")
```

### Which 5-minute interval contains the maximum number of steps?

```{r, results='asis'}
index <- which.max(interval$Mean)
print(xtable(interval[index,]), type = "html")

```

## Imputing missing values

### Calculating the total number of missig values

```{r}
sum(is.na(activity$steps))
```
We will apply is.na function to activity data set to remove NA values.

### Create a new data set and make a histgram of the total number of steps (imputing missing values)

```{r}
tolalNA <- sum(is.na(activity$steps))
bads <- is.na(activity$steps)
newdata <- activity[!bads,]

newsteps <- newdata %>% group_by(date) %>% summarise( Totalsteps = sum(steps), Mean = mean(steps), Median = median(steps))

barplot(newsteps$Totalsteps, space = 0, names.arg = newsteps$date,
        xlab = "Date",
        ylab = "Number of steps",
        main = "Total number of steps (imputing missing values)")
```

### Mean and median, total number of steps taken per day

```{r, results='asis'}

newstepsxt <- transform(newsteps, date = as.character(steps$date))
xt2 <- xtable(newstepsxt)
print(xt2, type = "html", include.rownames = FALSE)
              
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}

Sys.setlocale("LC_ALL","English")
week <- weekdays(newdata$date, abbreviate = TRUE)
weektype <- factor(week, 
                   levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"),
                   labels = c(rep("Weekday", 5), rep("Weelend", 2)))

newdataupdated <- newdata %>% mutate(weektype = weektype)
newstepsupdated <- newdataupdated %>% group_by(weektype, interval) %>% summarise(Mean = mean(steps))

ggplot(newstepsupdated, aes(x = interval, y = Mean)) +
  geom_line(col = "blue") +
  facet_wrap(~ weektype, nrow = 2) +
  labs(x = "Interval", y = "Number of steps", 
       title = "Activity patterns between weekdays and weekends")


```
